---
title: "EDA Budweiser dataset"
author: "jjsmu & rm"
date: "10/8/2023"
output: html_document
---

# Introduction
[See Summary memo](https://github.com/jjsmu/group16/tree/main#summary-memo-analysis-of-breweries-and-beer-data)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

sessionInfo()

# R libraries
#install.packages(dplyr)
library(dplyr)
#install.packages(ggplot)
library(ggplot2)
#install.packages("tidyverse")
library(tidyverse)
```

# Above Code: Installing R libraries needed for analysis 

# PART I Import data

```{r}
# read the csv files

BEERS = read.csv ("Beers.csv", header = TRUE)
BREWERIES = read.csv ("Breweries.csv", header = TRUE)
# check first five rows
head(BEERS, n = 5)
head(BREWERIES, n = 5)
# look at data frame
str(BEERS)
str(BREWERIES)
```

# Above Code: Checking Beer & breweries datasets before anlalysis. 


# PART II EDA Visualization
# 1.	How many breweries are present in each state?
# (See state_counts.csv)
```{r}
# visualization

# display summary statistics
summary(BEERS)
summary(BREWERIES)

# full table view
View(BEERS)
View(BREWERIES)

# view count of breweries by state
breweries_by_state <- BREWERIES %>% 
  group_by (State) %>%
  summarize (Count = n()) %>%
  arrange (desc(Count))
View(breweries_by_state)

# total number of breweries in US unfiltered
total_breweries <- sum(breweries_by_state$Count)
# total number of beers in US unfiltered
total_beers <- count(BEERS)
na_abv <- sum(is.na(BEERS$ABV))
na_abv
which(is.na(BEERS$ABV))
na_ibu <- sum(is.na(BEERS$IBU))
na_ibu

# ggplot2 bar plot of breweries by state
barplot_of_breweries_by_state <- BREWERIES %>%
  group_by(State) %>%
  ggplot(aes(State, fill = State)) + geom_bar(stat = "count") +
  ggtitle(paste("Barplot of Breweries by State (US), n = ", total_breweries)) +
  xlab("State") + ylab("Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
barplot_of_breweries_by_state

# TODO histogram of popular sizes by style
#hist(BEERS$Ounces, main = paste("Histogram of Beers by Volume (Ounce), n = ", length(table(beerOunceAsFactor))), xlab = "Volume (oz.)", ylab = "Frequency", col = "lightblue", border = "black")

plot(BEERS$IBU, BEERS$ABV, pch = 3, xlab = "IBU", ylab = "ABV(%)", main = paste("Scatter Plot of International Bitterness Unit v. Alcohol by Content, n = ", total_beers))
  
```

##  bar plot of breweries by state

```{r state, echo=FALSE}
barplot_of_breweries_by_state <- BREWERIES %>%
  group_by(State) %>%
  ggplot(aes(State, fill = State)) + geom_bar(stat = "count") +
  ggtitle(paste("Barplot of Breweries by State (US), n = ", total_breweries)) +
  xlab("State") + ylab("Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
barplot_of_breweries_by_state
```

# Above Code: We ran code to group and count breweries by state. [See State Count csv file](https://github.com/jjsmu/group16/blob/main/state_counts.csv)


# PART II
# 2. 	Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)
# First 6 Observations
```{r}
# full merge beer and breweries data frames 
BEER_AND_BREWERIES <- merge(BEERS, BREWERIES, by.x = "Brewery_id", by.y = "Brew_ID", all = TRUE) %>%
  rename(Beer = Name.x, Brewery = Name.y)
# view complete table
View(BEER_AND_BREWERIES)
# print the first 6 observations
print(head(BEER_AND_BREWERIES, n = 6))
# print the last 6 observations
print(tail(BEER_AND_BREWERIES, n = 6))
```

# Above Code: we checked the merged datasets


# PART III
# 3.	Address the missing values in each column.
```{r}

# tidy data
library(naniar)

# plot missing variables
gg_miss_var(BEERS) # IBU, ABV Missing
gg_miss_var(BREWERIES) 
gg_miss_var(BEER_AND_BREWERIES)

# test for NA values
sum(is.na(BEERS$ABV)) # missing 62 values
sum(is.na(BEERS$IBU)) # missing 1005 values
sum(is.na(BEER_AND_BREWERIES)) # missing 1067 values

# store index of rows containing NA
missing_index <- which(is.na(BEER_AND_BREWERIES))
missing_index

# test the function complete.cases() on gg_miss_var
gg_miss_var(BEERS[complete.cases(BEERS$IBU, BEERS$ABV), ])

# replace missing NA values with new value
# BEERS <- BEERS[complete.cases(BEERS$IBU, BEERS$ABV), ]

# NOTE: South Dakota IBU is NA for all, missing IBU data
```

# Above Code: Adressed missing values in datasets. See Codebook for additional information

# PART IV 
# 4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

```{r}

# filter and summarize ABV by state
median_by_state <- BEER_AND_BREWERIES %>%
  group_by(State) %>%
  #filter(!is.na(ABV) & !is.na(IBU)) %>%
  summarize(median_ABV = median(ABV, na.rm = TRUE), median_IBU = median(IBU, na.rm = TRUE))
(median_by_state)

# store sample size
total_states <- count(median_by_state)
total_states

# bar plot median ABV
barplot_median_abv <- median_by_state %>%
  ggplot(aes(State, median_ABV, color = median_ABV, fill = median_ABV)) + 
  geom_bar(stat = "identity") +
  ggtitle(paste("Barplot of Median Alcohol by Volume (% ABV) by State (US), n = ", total_states)) +
  xlab("State") + ylab("Median ABV (%)") + ylim(0, 0.15) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
barplot_median_abv

# bar plot median IBU
barplot_median_ibu <- median_by_state %>%
  ggplot(aes(State, median_IBU, color = median_IBU, fill = median_IBU)) +
  geom_bar(stat = "identity") + 
  ggtitle(paste("Barplot of Median Internation Bitterness Unit (IBU) by State (US), n = ", total_states)) +
  xlab("State(") + ylab("Median IBU") + ylim(0, 70) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
barplot_median_ibu

```
# PART V
#5.	Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r}

# group filter and summarize max abv and ibu values
max_by_state <- BEER_AND_BREWERIES %>%
  group_by(State) %>%
  #filter(!is.na(ABV) & !is.na(IBU)) %>%
  summarize(max_ABV = max(ABV, na.rm = TRUE), max_IBU = max(IBU, na.rm = TRUE))
max_by_state  

# bar plot max ABV 
barplot_max_abv <- max_by_state %>%
  ggplot(aes(State, max_ABV, color = max_ABV, fill = max_ABV)) +
  geom_bar(stat = "identity") + 
  ggtitle(paste("Barplot of Maximum Alcohol by Volume (% ABV) by State (US), n = ", total_states)) +
  xlab("State") + ylab("Max ABV (%)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
barplot_max_abv

# bar plot max IBU
barplot_max_ibu <- max_by_state %>%
  ggplot(aes(State, max_IBU, color = max_IBU, fill = max_IBU)) +
  geom_bar(stat = "identity") + 
  ggtitle(paste("Barplot of Maximum Internation Bitterness Unit (IBU) by State (US), n = ", total_states)) +
  xlab("State") + ylab("Maximum IBU") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
barplot_max_ibu

# console output max abv and ibu values
max(max_by_state$max_ABV)
max(max_by_state$max_IBU)

# store max abv and ibu index
max_abv <- which.max(max_by_state$max_ABV)
max_ibu <- which.max(max_by_state$max_IBU)

# console output max abv and ibu pointers
max_abv
max_ibu

# find max abv and ibu state by index
max_abv_state <- max_by_state$State[max_abv]
max_ibu_state <- max_by_state$State[max_ibu]

# console output max abv and ibu
max_abv_state # "CO"
max_ibu_state # "OR"

# import map libraries
library(maps)
library(mapproj)

# merge max and med abv and ibu
max_med_abv_ibu <- merge(max_by_state, median_by_state, by = "State", all = TRUE)
max_med_abv_ibu

# merge brewery count with max and med abv and ibu
breweries_map_data <- merge(max_med_abv_ibu, breweries_by_state, by = "State")
#colnames(breweries_map_data)[1] = "abb"
breweries_map_data$State <- trimws(breweries_map_data$State)
breweries_map_data

# store state in lookup dataframe
lookup = data.frame(State = state.abb, region = state.name) #makes a data frame with State name and abbreviation. 
lookup <- bind_rows(lookup, data.frame(State = "DC", region = "Washington DC")) # Add missing row
lookup <- lookup %>% arrange(lookup$State)
lookup$region <- tolower(lookup$region)
#lookup

# create state map data frame
us_states <- map_data("state")
us_states$subregion <- NULL
#us_states

# Found the bug! Blank space prevents merging, use trimws function
unique(breweries_map_data$State)
unique(lookup$State)

# merge breweries and lookup data 
breweries_lookup <- merge(breweries_map_data, lookup, by = "State")
breweries_lookup

# merge breweries lookup and us states by region
map.df <- merge(breweries_lookup, us_states, by = "region", all.x = T)
map.df <- map.df[order(map.df$order),] # arrange order desc, fix polygon clipping
#map.df

# plot the map data
breweries_heat_map <- map.df %>% ggplot(aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill = Count))+
  geom_path() + 
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+ggtitle(paste("Number of Breweries by State, n = ", total_breweries)) +
coord_map()
breweries_heat_map
# max abv heat map
max_abv_heat_map <- map.df %>% ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = max_ABV)) + geom_path() + scale_fill_gradientn(colors = rev(heat.colors(10)), na.value = "grey90") +
  ggtitle(paste("Maximum Alcohol by Volume (% ABV) by State, n = ", total_states)) +
  coord_map()
max_abv_heat_map
# max ibu heat map
max_ibu_heat_map <- map.df %>% ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = max_IBU)) + geom_path() + scale_fill_gradientn(colors = rev(heat.colors(10)), na.value = "grey90") +
  ggtitle(paste("Maximum International Bitterness Unit (IBU) by State, n = ", total_states)) + 
  coord_map()
max_ibu_heat_map
# median abv heat map
med_abv_heat_map <- map.df %>% ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = median_ABV)) + geom_path() + scale_fill_gradientn(colors = rev(heat.colors(10)), na.value = "grey90") +
  ggtitle(paste("Median Alcohol by Volume (% ABV) by State, n = ", total_states)) +
  coord_map()
med_abv_heat_map
# median ibu heat map
med_ibu_heat_map <- map.df %>% ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = median_IBU)) + geom_path() + scale_fill_gradientn(colors = rev(heat.colors(10)), na.value = "grey90") +
  ggtitle(paste("Median International Bitterness Unit (IBU) by State, n = ", total_states)) +
  coord_map()
med_ibu_heat_map

# string_r, which, grepl
# TODO search for keywords 'Ale', 'Pilsner', 'Stout'
```

# Above code: CO has the maximum alcoholic (ABV) beer. OR has the most bitter (IBU) beer


# PART VI
# 6.	Comment on the summary statistics and distribution of the ABV variable.
```{r}

# Comments:
summary(BEER_AND_BREWERIES)

plot_breweries <- BEER_AND_BREWERIES %>%
  ggplot(aes(ABV, IBU, alpha = ABV, size = Ounces, color = IBU), na.rm = TRUE) + geom_point() + geom_jitter() +
  ggtitle(paste("Scatter Plot of Alcohol by Content (% ABV) v. International Bitterness Units (IBU) by State (US), n = ", total_beers)) + xlab("ABV") + ylab("IBU")
plot_breweries

total_beers <- count(BEER_AND_BREWERIES)
hist_abv <- BEER_AND_BREWERIES %>%
  select(Beer, ABV, Style, Ounces, Brewery) %>%
  #filter(!is.na(ABV)) %>%
  ggplot(aes(ABV), na.rm = TRUE) + geom_histogram(stat = "bin", bins = 50) +
  ggtitle(paste("Histogram of Alcohol by Volume (% ABV), n = ", total_beers)) +
  xlab("ABV") + ylab("Frequency")
hist_abv

hist_ibu <- BEER_AND_BREWERIES %>%
  select(Beer, IBU, Style, Ounces, Brewery) %>%
  #filter(!is.na(IBU)) %>%
  ggplot(aes(IBU), na.rm = TRUE) + geom_histogram(stat = "bin", bins = 50) +
  ggtitle(paste("Histogram of International Bitterness Units (IBU), n = ", total_beers)) +
  xlab("IBU") + ylab("Frequency")
hist_ibu

# box plot of ABV by Beer Style
boxplot(BEERS$ABV~BEERS$Style, las = 2, names.arg = BEERS$Style, main = paste("Boxplot of Alcohol by Volume by Style, n = ", length(BEERS$Style)), xlab = "Beer Style", ylab = "ABV(%)")
# box plot of IBU by Beer Style
boxplot(BEERS$IBU~BEERS$Style, las = 2, names.arg = BEERS$Style, main = paste("Boxplot of International Bitterness Units (IBU) by Style, n = ", length(BEERS$Style)), xlab = "Beer Style", ylab = "IBU")

```


# Above code: ABV is weakly correletd with higher IBU


# PART VII
# 7.	Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.
# Our observations suggest a weak positive correlation between the bitterness of the beer and its alcoholic content. A linear regression plot confirms this with IBU values increasing with respect to ABV range. The rate at which ABV increases as IBU increases is also greater with respect to IPA style beers. Furthermore, EDA of distributions of ABV by IBU in histogram plots show a center of mean of ABV and IBU for IPA greater than Ale beers. We conclude the relationship between bitterness and alcohol content corresponds to the "hoppiness" in beers.
```{r}

# scatter plot of IBU v. ABV
ggplot(data = BEER_AND_BREWERIES) +
  geom_point(mapping = aes(IBU, ABV, color = State, size = Ounces)) +
  geom_smooth(mapping = aes(IBU, ABV, linetype = as.factor(Ounces))) +
  ggtitle(paste("Scatter plot Alcohol by Volume (% ABV) v. International Bitterness Unit (IBU) by Ounce (oz.), n = ", total_beers))+
  xlab("ABV (%)") + ylab("IBU")  
  #facet_wrap(~Ounces)
```

# Above Code: There is an apparent relationship between the bitterness of the beer and its alcoholic content, the higher the IBU the more likely it is the the ABV is also higher

# PART VIII
#8.	Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN classification to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages … KNN is very easy to understand conceptually. 
#In addition, while you have decided to use KNN to investigate this relationship (KNN is required) you may also feel free to supplement your response to this question with any other methods or techniques you have learned.  Creativity and alternative solutions are always encouraged.  

```{r}
# filter by ale
ale_abv_ibu <- BEER_AND_BREWERIES %>%
  select(Ounces, Style, ABV, IBU) %>%
  filter(grepl("Ale", Style))
ale_abv_ibu$Class <- "Ale"
dim(ale_abv_ibu)
#ale_abv_ibu

# filter by ipa
ipa_abv_ibu <- BEER_AND_BREWERIES %>%
  select(Ounces, Style, ABV, IBU) %>%
  filter(grepl("IPA", Style))
ipa_abv_ibu$Class <- "IPA"

# combine ale and ipa list
ale_and_ipa <- rbind(ipa_abv_ibu, ale_abv_ibu)
```
## Plot ale and ipa distributions
```{r}
#install package and load library
library(GGally)

# plot ggpairs by class ipa v. ale
ggpairs_ale_ipa_abv_ibu_by_class <- ale_and_ipa %>%
  select(ABV, IBU, Class) %>% 
  ggpairs(aes(color = Class))
ggpairs_ale_ipa_abv_ibu_by_class
# suppress warning: standard deviation is zero - ABV, IBU NA values
```
## KNN Classification
# Evaluate the ML classification model using K-Nearest Neighbor algorithm to predict ale and IPA class
```{r}
# install KNN libraries
#install.packages("class")
#install.packages("caret")
#install.packages("e1071")
library(class)
library(caret)
library(e1071)

# count total ale and ipa
count(ale_and_ipa) # 1547 samples total, including NA
ale_and_ipa <- na.omit(ale_and_ipa) # omit NA values
count(ale_and_ipa) # omit NA, 951 remain

# set up train and test set
trainIndex <- sample(seq(1, 951), 665)
trainBeers <- ale_and_ipa[trainIndex,]
testBeers <- ale_and_ipa[-trainIndex,]

# plot train set
trainBeers %>% ggplot(aes(ABV, IBU, color = Class)) + geom_point()

# knn classification
classification = knn(trainBeers[, c("ABV", "IBU")], testBeers[, c("ABV", "IBU")], trainBeers$Class, k = 5, prob = TRUE )
classification
table <- table(classification, testBeers$Class)
table
CM <- confusionMatrix(table, mode = "everything")
CM
```
## Internal LOOCV
# Use leave one out cross validation function to estimate ale and ipa classification model performance
```{r}
# k = 5, train test only
loocv_classification <- knn.cv(trainBeers[, c("ABV", "IBU")], trainBeers$Class, k = 5)
loocv_classification
loocv_table <- table(loocv_classification, trainBeers$Class)
loocv_table
loocv_cm <- confusionMatrix(loocv_table, mode = "everything")
loocv_cm
```

Above code: KNN support the finding that IBU and ABV seem to be positivly correlated  

# PART IX
#9.	Knock their socks off!  Find one other useful inference from the data that you feel Budweiser may be able to find value in.  You must convince them why it is important and back up your conviction with appropriate statistical evidence. 

```{r}
# plot ggpairs by ounce ipa v. ale
ggpairs_ale_ipa_abv_ibu_by_ounces <- ale_and_ipa %>%
  select(ABV, IBU, Ounces) %>% 
  ggpairs(aes(na.rm = TRUE, color = as.factor(Ounces)))
ggpairs_ale_ipa_abv_ibu_by_ounces

```

Above code: EDA distribution plot of all beers by Ounce. The 3x3 plot shows the distribution of ABV and IBU values of beers by serving size. There is no evidence to suggest a correlation between ABV and IBU values with respect to ounce, however shows that 12 and 16 oz. servings are the most popular sizes. With additional data such as pricing, this information could be used, for example, to determine an entry point for Budweiser to compare and price their new products with respect to other distributors in the US.




