---
title: "EDA Budweiser dataset"
author: "jjsmu"
date: "10/8/2023"
output: html_document

#1.	How many breweries are present in each state?
#2.	Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)
#3.	Address the missing values in each column.
#4.	Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
#5.	Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
#6.	Comment on the summary statistics and distribution of the ABV variable.
#7.	Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.
#8.	Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN classification to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages … KNN is very easy to understand conceptually. 
#In addition, while you have decided to use KNN to investigate this relationship (KNN is required) you may also feel free to supplement your response to this question with any other methods or techniques you have learned.  Creativity and alternative solutions are always encouraged.  

#9.	Knock their socks off!  Find one other useful inference from the data that you feel Budweiser may be able to find value in.  You must convince them why it is important and back up your conviction with appropriate statistical evidence. 

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# R libraries
#install.packages(dplyr)
library(dplyr)
#install.packages(ggplot)
library(ggplot2)
#install.packages("tidyverse")
library(tidyverse)
```
# PART I Import data

```{r}
# read the csv files
BEERS = read.csv(file.choose(), header = TRUE)
BREWERIES = read.csv(file.choose(), header = TRUE)
# check first five rows
head(BEERS, n = 5)
head(BREWERIES, n = 5)
# look at data frame
str(BEERS)
str(BREWERIES)
```

# PART II EDA Visualization
# 1.	How many breweries are present in each state?
```{r}
# visualization

# display summary statistics
summary(BEERS)
summary(BREWERIES)

# full table view
View(BEERS)
View(BREWERIES)

# view count of breweries by state
breweries_by_state <- BREWERIES %>% 
  group_by (State) %>%
  summarize (Count = n()) %>%
  arrange (desc(Count))
View(breweries_by_state)

# total number of breweries in US unfiltered
total_breweries <- sum(breweries_by_state$Count)
# total number of beers in US unfiltered
total_beers <- count(BEERS)

# ggplot2 bar plot of breweries by state
barplot_of_breweries_by_state <- BREWERIES %>%
  group_by(State) %>%
  ggplot(aes(State, fill = State)) + geom_bar(stat = "count") +
  ggtitle(paste("Barplot of Breweries by State (US), n = ", total_breweries)) +
  xlab("State") + ylab("Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
barplot_of_breweries_by_state

# TODO histogram of popular sizes by style
#hist(BEERS$Ounces, main = paste("Histogram of Beers by Volume (Ounce), n = ", length(table(beerOunceAsFactor))), xlab = "Volume (oz.)", ylab = "Frequency", col = "lightblue", border = "black")

plot(BEERS$IBU, BEERS$ABV, pch = 3, xlab = "IBU", ylab = "ABV(%)", main = paste("Scatter Plot of International Bitterness Unit v. Alcohol by Content, n = ", total_beers))
  
```
# PART II
# 2. 	Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)
```{r}
# full merge beer and breweries data frames 
BEER_AND_BREWERIES <- merge(BEERS, BREWERIES, by.x = "Brewery_id", by.y = "Brew_ID", all = TRUE) %>%
  rename(Beer = Name.x, Brewery = Name.y)
# view complete table
View(BEER_AND_BREWERIES)
# print the first 6 observations
print(head(BEER_AND_BREWERIES, n = 6))
# print the last 6 observations
print(tail(BEER_AND_BREWERIES, n = 6))
```

# PART III
# 3.	Address the missing values in each column.
```{r}

# tidy data
library(naniar)

# plot missing variables
gg_miss_var(BEERS) # IBU, ABV Missing
gg_miss_var(BREWERIES) 
gg_miss_var(BEER_AND_BREWERIES)

# test for NA values
sum(is.na(BEERS$ABV)) # missing 62 values
sum(is.na(BEERS$IBU)) # missing 1005 values
sum(is.na(BEER_AND_BREWERIES)) # missing 1067 values

# TODO find the missing NA value indexes
missing_index <- which(grep(NA, BEER_AND_BREWERIES))
missing_index

# test the function complete.cases() on gg_miss_var
gg_miss_var(BEERS[complete.cases(BEERS$IBU, BEERS$ABV), ])

# replace missing NA values with new value
# BEERS <- BEERS[complete.cases(BEERS$IBU, BEERS$ABV), ]

# NOTE: South Dakota IBU is NA for all, missing IBU data
```
# PART IV 
# 4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

```{r}

# filter and summarize ABV by state
median_by_state <- BEER_AND_BREWERIES %>%
  group_by(State) %>%
  #filter(!is.na(ABV) & !is.na(IBU)) %>%
  summarize(median_ABV = median(ABV, na.rm = TRUE), median_IBU = median(IBU, na.rm = TRUE))
median_by_state

# store sample size
total_states <- count(median_by_state)
total_states

# bar plot median ABV
barplot_median_abv <- median_by_state %>%
  ggplot(aes(State, median_ABV, color = median_ABV, fill = median_ABV)) + 
  geom_bar(stat = "identity") +
  ggtitle(paste("Barplot of Median Alcohol by Volume (% ABV) by State (US), n = ", total_states)) +
  xlab("State") + ylab("Median ABV (%)") + ylim(0, 0.15) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
barplot_median_abv

# bar plot median IBU
barplot_median_ibu <- median_by_state %>%
  ggplot(aes(State, median_IBU, color = median_IBU, fill = median_IBU)) +
  geom_bar(stat = "identity") + 
  ggtitle(paste("Barplot of Median Internation Bitterness Unit (IBU) by State (US), n = ", total_states)) +
  xlab("State(") + ylab("Median IBU") + ylim(0, 70) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
barplot_median_ibu

```
# PART V
#5.	Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r}

# group filter and summarize max abv and ibu values
max_by_state <- BEER_AND_BREWERIES %>%
  group_by(State) %>%
  #filter(!is.na(ABV) & !is.na(IBU)) %>%
  summarize(max_ABV = max(ABV, na.rm = TRUE), max_IBU = max(IBU, na.rm = TRUE))
max_by_state  

# bar plot max ABV 
barplot_max_abv <- max_by_state %>%
  ggplot(aes(State, max_ABV, color = max_ABV, fill = max_ABV)) +
  geom_bar(stat = "identity") + 
  ggtitle(paste("Barplot of Maximum Alcohol by Volume (% ABV) by State (US), n = ", total_states)) +
  xlab("State") + ylab("Max ABV (%)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
barplot_max_abv

# bar plot max IBU
barplot_max_ibu <- max_by_state %>%
  ggplot(aes(State, max_IBU, color = max_IBU, fill = max_IBU)) +
  geom_bar(stat = "identity") + 
  ggtitle(paste("Barplot of Maximum Internation Bitterness Unit (IBU) by State (US), n = ", total_states)) +
  xlab("State") + ylab("Maximum IBU") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
barplot_max_ibu

# console output max abv and ibu values
max(max_by_state$max_ABV)
max(max_by_state$max_IBU)

# store max abv and ibu index
max_abv <- which.max(max_by_state$max_ABV)
max_ibu <- which.max(max_by_state$max_IBU)

# console output max abv and ibu pointers
max_abv
max_ibu

# find max abv and ibu state by index
max_abv_state <- max_by_state$State[max_abv]
max_ibu_state <- max_by_state$State[max_ibu]

# console output max abv and ibu
max_abv_state # "CO"
max_ibu_state # "OR"

# import map libraries
library(maps)
library(mapproj)

# merge max and med abv and ibu
max_med_abv_ibu <- merge(max_by_state, median_by_state, by = "State", all = TRUE)
max_med_abv_ibu

# merge brewery count with max and med abv and ibu
breweries_map_data <- merge(max_med_abv_ibu, breweries_by_state, by = "State")
#colnames(breweries_map_data)[1] = "abb"
breweries_map_data$State <- trimws(breweries_map_data$State)
breweries_map_data

# store state in lookup dataframe
lookup = data.frame(State = state.abb, region = state.name) #makes a data frame with State name and abbreviation. 
lookup <- bind_rows(lookup, data.frame(State = "DC", region = "Washington DC")) # Add missing row
lookup <- lookup %>% arrange(lookup$State)
lookup$region <- tolower(lookup$region)
lookup

# create state map data frame
us_states <- map_data("state")
us_states$subregion <- NULL
us_states

# Found the bug! Blank space prevents merging, use trimws function
unique(breweries_map_data$State)
unique(lookup$State)

# merge breweries and lookup data 
breweries_lookup <- merge(breweries_map_data, lookup, by = "State")
breweries_lookup

# merge breweries lookup and us states by region
map.df <- merge(breweries_lookup, us_states, by = "region", all.x = T)
map.df <- map.df[order(map.df$order),] # arrange order desc, fix polygon clipping
map.df

# plot the map data
breweries_heat_map <- map.df %>% ggplot(aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill = Count))+
  geom_path() + 
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+ggtitle(paste("Number of Breweries by State, n = ", total_breweries)) +
coord_map()
breweries_heat_map
# max abv heat map
max_abv_heat_map <- map.df %>% ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = max_ABV)) + geom_path() + scale_fill_gradientn(colors = rev(heat.colors(10)), na.value = "grey90") +
  ggtitle(paste("Maximum Alcohol by Volume (% ABV) by State, n = ", total_states)) +
  coord_map()
max_abv_heat_map
# max ibu heat map
max_ibu_heat_map <- map.df %>% ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = max_IBU)) + geom_path() + scale_fill_gradientn(colors = rev(heat.colors(10)), na.value = "grey90") +
  ggtitle(paste("Maximum International Bitterness Unit (IBU) by State, n = ", total_states)) + 
  coord_map()
max_ibu_heat_map
# median abv heat map
med_abv_heat_map <- map.df %>% ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = median_ABV)) + geom_path() + scale_fill_gradientn(colors = rev(heat.colors(10)), na.value = "grey90") +
  ggtitle(paste("Median Alcohol by Volume (% ABV) by State, n = ", total_states)) +
  coord_map()
med_abv_heat_map
# median ibu heat map
med_ibu_heat_map <- map.df %>% ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = median_IBU)) + geom_path() + scale_fill_gradientn(colors = rev(heat.colors(10)), na.value = "grey90") +
  ggtitle(paste("Median International Bitterness Unit (IBU) by State, n = ", total_states)) +
  coord_map()
med_ibu_heat_map

# string_r, which, grepl
# TODO search for keywords 'Ale', 'Pilsner', 'Stout'
```

# PART VI
# 6.	Comment on the summary statistics and distribution of the ABV variable.
```{r}

# Comments:
summary(BEER_AND_BREWERIES)

plot_breweries <- BEER_AND_BREWERIES %>%
  ggplot(aes(ABV, IBU)) + geom_point() %>%
  ggtitle(paste("Scatter Plot of Alcohol by Content v. International Bitterness Units, n = ", total_beers )) + xlab("ABV") + ylab("IBU")
plot_breweries

total_beers <- count(BEER_AND_BREWERIES)
hist_abv <- BEER_AND_BREWERIES %>%
  select(Beer, ABV, Style, Ounces, Brewery) %>%
  filter(!is.na(ABV)) %>%
  ggplot(aes(ABV)) + geom_histogram(bins = 30, color = "red", fill = "pink") +
  ggtitle(paste("Histogram of Alcohol by Volume (% ABV), n = ", total_beers)) +
  xlab("ABV") + ylab("Frequency")
hist_abv

hist_ibu <- BEER_AND_BREWERIES %>%
  select(Beer, IBU, Style, Ounces, Brewery) %>%
  filter(!is.na(IBU)) %>%
  ggplot(aes(IBU)) + geom_histogram(stat = "bin", bins = 50) +
  ggtitle(paste("Histogram of International Bitterness Units (IBU), n = ", total_beers)) +
  xlab("IBU") + ylab("Frequency")
hist_ibu

# box plot of ABV by Beer Style
boxplot(BEERS$ABV~BEERS$Style, las = 2, names.arg = BEERS$Style, main = paste("Boxplot of Alcohol by Volume by Style, n = ", length(BEERS$Style)), xlab = "Beer Style", ylab = "ABV(%)")
# box plot of IBU by Beer Style
boxplot(BEERS$IBU~BEERS$Style, las = 2, names.arg = BEERS$Style, main = paste("Boxplot of International Bitterness Units (IBU) by Style, n = ", length(BEERS$Style)), xlab = "Beer Style", ylab = "IBU")

```

# PART VII
# 7.	Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.
```{r}

# scatter plot of IBU v. ABV
ggplot(data = BEER_AND_BREWERIES) +
  geom_point(mapping = aes(IBU, ABV, color = State)) +
  geom_smooth(mapping = aes(IBU, ABV, linetype = as.factor(Ounces))) +
  ggtitle(paste("Scatter plot Alcohol by Volume (% ABV) v. International Bitterness Unit (IBU) by Ounce (oz.), n = ", total_beers))+
  xlab("ABV (%)") + ylab("IBU") +
  facet_wrap(~Ounces)

gg
```

# PART VIII
#8.	Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN classification to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages … KNN is very easy to understand conceptually. 
#In addition, while you have decided to use KNN to investigate this relationship (KNN is required) you may also feel free to supplement your response to this question with any other methods or techniques you have learned.  Creativity and alternative solutions are always encouraged.  

```{r}
# filter for non null ale beers
ale_abv_ibu <- BEER_AND_BREWERIES %>%
  select(Ounces, Style, ABV, IBU) %>%
  filter(!is.na(ABV) & !is.na(IBU) & grepl("Ale", Style)) # without NA
  #filter(grepl("Ale", Style))
ale_abv_ibu$Class <- "Ale"
ale_abv_ibu

# filter for non null IPA beers
ipa_abv_ibu <- BEER_AND_BREWERIES %>%
  select(Ounces, Style, ABV, IBU) %>%
  filter(!is.na(ABV) & !is.na(IBU) & grepl("IPA", Style)) # without NA
  #filter(grepl("IPA", Style))
ipa_abv_ibu$Class <- "IPA"
ipa_abv_ibu

# combine ale and ipa list
ale_and_ipa <- rbind(ipa_abv_ibu, ale_abv_ibu)
ale_and_ipa

#install package and load library
library(GGally)

# group by ale and ipa
ggpairs_ale_ipa_abv_ibu <- ale_and_ipa %>%
  select(Ounces, ABV, IBU, Class) %>% 
  mutate(ouncesAsFactor = as.factor(Ounces)) %>%
  group_by(Class) %>%
  ggpairs(aes(color = ouncesAsFactor))
ggpairs_ale_ipa_abv_ibu

# TODO mutate new column IPA or Ale group by style

install.packages("class")
install.packages("caret")
install.packages("e1071")
library(class)
library(caret)
library(e1071)

# KNN 
train <- ale_and_ipa
test <- data.frame(ABV = 0.08, IBU = 60)
classification = knn(train[,3:4], test, train$Class, k = 5, prob = TRUE)
matrix = confusionMatrix(table(classification, test$Class)) # TODO fix me

# KNN cross validation

count(ale_and_ipa) # 951 samples total
# set up train and test set
trainIndex <- sample(seq(1, count(ale_and_ipa), 666))
trainBeers <- ale_and_ipa[trainIndex,]
testBeers <- ale_and_ipa[-trainIndex,]

# plot train set
trainBeers %>%

```

# PART IX
#9.	Knock their socks off!  Find one other useful inference from the data that you feel Budweiser may be able to find value in.  You must convince them why it is important and back up your conviction with appropriate statistical evidence. 

```{r}
# use of themes and advanced plots
hist_abv <- BEER_AND_BREWERIES %>%
  ggplot(aes(ABV, na.rm = TRUE)) + geom_histogram()
hist_abv
#TODO bar plot stack breweries by state
#TODO bar plot stack ounces by state
#TODO bar plot stack "Class" by state

#TODO ggplotly
#install.packages(ggthemes)
#install.packages(ggplotly)
library(ggthemes)
library(ggplotly)

#eda 1: number of breweries v. median ibu
```

